\section{Challenges and Future Work}\label{sec:conclusions}
In general, our development experience is mirrored by the lessons described 
in~\cite{minao-hspr18} and ~\cite{bertin-netdev17}.

\paragraph{No multi-/broadcast support}
While XDP is able to redirect single frames it does not have the ability to 
clone and redirect multiple packets similar to \texttt{bpf\_clone\_redirect}. 
This makes development of more sophisticated P4 forwarding programs problematic.

\paragraph{The stack size is too small}
More complex generated XDP programs are rejected by the verifier despite their 
safeness. 
This is a particular challenge when attempting to implement network function 
chaining or advanced pipelined packet processing in a single XDP program. 

\paragraph{Generic XDP and TCP}
Our testing framework uses virtual Linux interfaces and generic 
XDP~\cite{genericxdp} to verify XDP programs. 
Unfortunately, we are unable to test TCP streams as the protocol is not 
supported by this driver~\cite{xdptcp}.
Any program loaded by generic XDP operates after the creation of
the \texttt{skb} and requires the original packet data. Since TCP clones every 
packet and passes the unmodifiable \texttt{skb} clone, generic XDP is
bypassed and never receives the datagram.

\paragraph{Using the libbpf userspace library}
Compilation of eBPF programs in user-space requires substantial 
effort. Many function calls and variables available in sample programs are not 
available as general C library. Any user trying to interact with the 
generated C code has to provide their own sources. Currently, P4C-XDP maintains 
copies of utilities from kernel code or various online sources. This is not a 
sustainable approach. We plan to integrate \texttt{libbpf} with our repository 
to control and manage the eBPF programs and maps.

\paragraph{Pinned eBPF maps in network namespaces}
When using eBPF programs in namespaces, maps that were pinned via \texttt{tc} 
do not persist across \texttt{ip netns exec} calls. As consequence, any 
program has to be run in a single continuous shell command. Example:
{\scriptsize
	\begin{verbatim}
bash -c "tc filter add ...; ls  /sys/fs/bpf/tc/globals"
	\end{verbatim}
}
Once \texttt{ip netns exec} has finished, the reference to the eBPF 
map and all its associated state disappear. The eBPF program, however, remains 
attached to the virtual interface, leading to inconsistent packet processing 
behavior.
